Relational Parametricity for Algebraically Indexed Types

### 2012-06-07 paper/section-introduction.tex
Substantial changes to the motivating examples section

@@

>>>>

Possible angles of attack:
\begin{itemize}
\item Parametric polymorphic types allow us to prevent
  over-specification of the behaviour of programs. For instance, the
  type $\forall \alpha. [\alpha] \to [\alpha]$ is a generalisation of
  the types $[\mathsf{int}] \to [\mathsf{int}]$ and $[\mathsf{char}]
  \to [\mathsf{char}]$. Either of the latter two types over-specify
  the behaviour of the function.
\item There are other cases of programs that are over-specified. The
  leading example we just below is of geometric programs that
  manipulate coordinate data. Often, programs that manipulate
  coordinate data are insensitive to geometric transformations. For
  example, a program that computes the area of a triangle described by
  three points is insensitive to translations or rotations applied to
  all three points.
\item 
\end{itemize}

@@

\begin{itemize}
\item Formulation of indexed parametricity that talks about
  terms, not as part of terms (FIXME)
\item Therefore, it can act as an analysis of programs, which can be
  imposed after the fact
\item Allow for more refined relational interpretations in order to
  derive non-definability results.
\item 
\end{itemize}

%%% Local Variables:
%%% TeX-master: "paper"
%%% End:

### 2012-06-07 paper/section-motivating-examples.tex
Add AUCTeX commands to deal with multifile document
Substantial changes to the motivating examples section

FIXME:
\begin{enumerate}
\item Describe the units of measure system
\item Describe the concept of scale invariance
\item Describe the consequences of scale invariance
\end{enumerate}

\subsection{Vector Spaces and Frame Invariance}
\label{sec:vector-spaces-intro}

@@

>>>>

Kennedy's units of measure system enforces, through a static typing
discipline, a semantic notion of unit correctness. Idea: units of
measure $\Rightarrow$ scale invariance $\Rightarrow$ representation
independence $\Rightarrow$ Reynolds-style parametricity.

To illustrate the semantic formulation of unit correctness, we present
a small example.  Let $f$ be a function that takes pairs of real
numbers to a real number:
\begin{displaymath}
  f : \mathbb{R} \times \mathbb{R} \to \mathbb{R},
\end{displaymath}
and let us assume that the first input to the function should denote
some quantity of distance, the second input should denote some
quantity of time, and that the output should denote some quantity of
distance over time. Intuitively, these unit constraints mean that $f$
may only do something like divide the first argument by the second,
and possibly multiply by some dimensionless constant. But how do we
state such conditions formally in terms of $f$?

One way to state unit correctness is in terms of preservation of
scaling. If we scale all distance inputs by some factor $k_d$, and all
time inputs by some factor $k_t$, then the output of $f$ should
correspondingly scale by the factor $k_d/k_t$ (since the supposed
units of the output is distance over time). In symbols:
\begin{displaymath}
  \forall k_d, k_t, x_d, x_t.\ f(k_dx_d, k_tx_t) = \frac{k_d}{k_t}f(x_d,x_t)
\end{displaymath}

Another way of looking at scale invariance is in terms of invariance
under change of representation. The number $5$, with respect to

\begin{enumerate}
\item Semantic correctness for units of measure is enforced by scaling
  invariance
\item Systematise this by using Reynolds' idea that semantic type
  abstraction is formulated as preservation of relations. In the units
  of measure case, we assign a relational meaning to every unit
  expression.
\item Consequences...
\end{enumerate}

\subsection{Affine Spaces, Vector Spaces and Frame Invariance}
\label{sec:vector-spaces-intro}

An affine space is a vector space that has ``forgotten its origin''.

@@

% \subsection{Monadic Effects}
% \label{sec:monadic-effects}

% Maybe this will work...
